A distance derived from the Kolmogorov-Smirnov statistic: specification, reference measures and example uses

### 2017-08-23 emc/article.tex
finished first draft

\begin{abstract}
Statistical distances quantifies the difference between two statistical constructs.
In this article, we describe reference values for a semimetric
derived from the Kolmogorov-Smirnov statistic $D_{F,F'}$.
Each measure of the $D_{F,F'}$ is a distance between two histograms.
This distance is normalized by the number of observations in each sample
to yield the $c'=D_{F,F'}\sqrt{\frac{n n'}{n+n'}}$ statistic,
which can be mapped to p-values, i.e. values for which 
high enough levels of significance $\alpha$ implies the rejection of the
	null hypothesis (that the samples are drawn from the same distribution).
One great feature of $c'$ is that it inherits the robustness of
	$D_{F,F'}$ and is thus suitable for use in settings where
	the underlying distributions are not known.
Benchmarks are derived by comparing samples from standard distributions.
The supplied example applications of the $c'$ statistic for the distinction
	of samples in real data enables further
insights about the robustness and power of $c$.
\end{abstract}

@@

\subsection{Description of the $c'$ statistic}
Be $F$ and $F'$ two empirical cumulative distributions,
where $n$ and $n'$ are the number of observations in each sample.
The two-sample Kolmogorov-Smirnov test rejects the null hypothesis,
that the histograms are the outcome of the same underlying distribution,
if:
\begin{equation}\label{eq:ks}
D_{F,F'} > c(\alpha)\sqrt{\frac{n+n'}{nn'}}
\end{equation}

@@

\subsubsection{Varying the sample sizes}
\subsection{Example application to real samples}
\subsection{A thorough exposition of the tables}
\subsection{Scripts for automated generation of the tables}

\section{CONCLUSIONS AND FURTHER WORK}\label{sec:conc}
% discrete vs continous

% are histograms really needed?

% make a better presentation of the benchmark tables

% organize the scripts

@@

% ------------------------------------------------------------------------
\begin{thebibliography}{99}
\fontsize{11}{0}\selectfont
\bibitem[DrV, 2016]{stack}
	DrV (2016). Python script to obtain the CDF from samples without doing the histogram.
		Stack Overflow.
		From \url{https://stackoverflow.com/questions/24788200/calculate-the-cumulative-distribution-function-cdf-in-python}

\bibitem[Fabbri, 2017]{enhance}
	Fabbri, R. (2017). Enhancements of linked data expressiveness for ontologies.
		Encontro Nacional de Modelagem Computacional 2017 (XX ENMC).
		From \url{https://github.com/ttm/ontologyEnhancements/raw/master/article.pdf}

\bibitem[Heath \& Bizer, 2011]{ldb}
	Heath, T. \& Bizer, C. (2011). Linked Data: Evolving the Web into a Global Data Space (1st edition). Synthesis Lectures on the Semantic Web: Theory and Technology, 1:1, 1-136. Morgan \& Claypool.

\bibitem[Lehmann et al., 2015]{dbpedia}
	Lehmann, J., Isele, R., Jakob, M., Jentzsch, A., Kontokostas, D., Mendes, P. N., ... \& Bizer, C. (2015). DBpedia–a large-scale, multilingual knowledge base extracted from Wikipedia. Semantic Web, 6(2), 167-195.

\bibitem[Munzner, 2014]{munzner}
	Munzner, T. (2014). Visualization analysis and design. CRC press.

\bibitem[W3C, 2010]{w3cld}
	W3C (2017). LINKED DATA CURRENT STATUS, from \url{https://www.w3.org/standards/techs/linkeddata}

\bibitem[Ward et al., 2010]{ward}
	Ward, M. O., Grinstein, G., \& Keim, D. (2010). Interactive data visualization: foundations, techniques, and applications. CRC Press.

\bibitem[Wikipedia, 2017]{wikipData}
	Data. (2017, August 21). In Wikipedia, The Free Encyclopedia. Retrieved
		22:31, August 21, 2017
		, from \url{https://en.wikipedia.org/w/index.php?title=Data&oldid=796493851}

\end{thebibliography}
% ------------------------------------------------------------------------

>>>>

\begin{abstract}
Statistical distances quantifies the difference between two statistical constructs.
In this article, we describe reference values for a semimetric
derived from the Kolmogorov-Smirnov statistic $D_{F,F'}$.
Each measure of the $D_{F,F'}$ is a distance between two histograms.
This distance is normalized by the number of observations in each sample
to yield the $c'=D_{F,F'}\sqrt{\frac{n n'}{n+n'}}$ statistic,
which can be mapped to p-values, i.e. values for which 
high enough levels of significance $\alpha$ implies the rejection of the
	null hypothesis (that the samples are drawn from the same distribution).
One great feature of $c'$ is that it inherits the robustness of
	$D_{F,F'}$ and is thus suitable for use in settings where
	the underlying distributions are not known.
Benchmarks are derived by comparing samples from standard distributions.
The supplied example applications of the $c'$ statistic for the distinction
	of samples in real data enables further
insights about the robustness and power of such statistical distance.
\end{abstract}

@@

\subsection{Description of the $c'$ statistic}\label{sec:desc}
Be $F$ and $F'$ two empirical cumulative distributions,
where $n$ and $n'$ are the number of observations in each sample.
The two-sample Kolmogorov-Smirnov test rejects the null hypothesis,
that the histograms are the outcome of the same underlying distribution,
if:
\begin{equation}\label{eq:ks}
D_{F,F'} > c(\alpha)\sqrt{\frac{n+n'}{nn'}}
\end{equation}

@@

\subsection{Example application to real samples}
To further validate the $c'$ statistic and enable deeper insights,
a number of applications to real samples were performed:
\begin{itemize}
	\item Texts: Hamlet (Shakespeare), the Bible (KJV), Moby Dick (Herman Melville) and Esaú e Jacó (Machado de Assis),
		where studied by regarding the stopwords and the words which were not stopwords.
		Each of these works were considered as a whole and divided in the first and last half.
                These texts were used to obtain samples that are:
		the mean of the token sizes, the standard variation of the token sizes,
		the token sizes.
		For the two first samples, the text was divided into 1000 parts in which the means
		and standard variations where obtained.
		The overall result is: smaller $c'$ for comparisons between parts of the same text
		although high $c'$ was incident even between parts of the same book.
	\item Audio: the audio segments for testing the sound system of a Ubuntu Linux distribution
		were considered both by their PCM samples and by their Daubechies 8 wavelet coefficients.
		The segments yielded higher $c'$ values as the audio were more different, e.g. the noise
		yielded the higher $c'$ when compared against the other signals.
	\item Music: each classical composition were regarded as a sample
		and the pitches were regarded as observations.
		The results reflect music history.
		For example, measures of $c$ involving Palestrina
		increases with the exception of Beethoven
		who, indeed, used modalism.
		The values of $c$ related to Bach also increases along time,
		and the outcome of the comparison against Palestrina 
		is only exceeded when Sch\"onberg is reached,
		which reflects the non-tonal discourse of both
		Palestrina and Sch\"onberg.
		Table~\ref{tab:mus} exposes these results.
	\item OS status: workload of the CPUs and of the most consuming processes.
		Again, the type of samples are mandatory:
		they might all be identified by the values of $c'$
		found in comparison to other samples,
		with the exception of the RAM memory.
\end{itemize}

\input{../tables/musicDistances_Foo}

\subsection{A thorough exposition of the tables}
These results yield many tables which do not fit this article and would make
this exposition clumsy.
Their thorough expositions are in~\cite{ksstats},
eich all the tables and descriptions.

\subsection{Scripts for automated generation of the tables}
The tables that are benchmarks and that result from comparing real samples
are rendered by scripts.
These scripts are configurable, i.e. might be set to render other tables if needed.
Once the new tables are rendered, they might be assembled into a PDF by means
of the latex files that yield~\cite{ksstats}.

\section{CONCLUSIONS AND FURTHER WORK}\label{sec:conc}
This exposition described the $c'$ statistic, which is a statistical distance,
the benchmark tables for $c'$ and its use to observe differences in real samples.
As far as I understand, the tables are effective in exposing reference
values in various settings of various distributions.
The Kolmogorv-Smirnov test, from which $c'$ is derived, is known
to be robust in the sense that it is usable even when the underlying distributions
are not known or present problems for other tests.
The overall result is that we obtained a statistical distance which is
useful in various contexts and have now benchmark tables.

Potential next steps are:
\begin{itemize}
	\item better organize the scripts that render the benchmark tables,
		because they are scattered in the \texttt{tests/} directory
		of~\cite{gmaneLegacy}.
	\item Make a better presentation of the benchmark tables in~\cite{ksstats}.
		They are sound but were made for personal usage and might be enhanced
		by enhancements in the writing of their descriptions and contextualization.
	\item Use other distributions for obtaining the tables.
		This is relevant mainly because the number of rejections of the null hypothesis
		was sometimes higher that expected for the significance level in power-law distributions.
	\item Compare $c'$ to other statistical distances: in which cases are they suitable and
		what results they yield.
\end{itemize}
\noindent Finally, the most urgent developments this contribution need are:
1) understand the differences in $c'$ in the case of continuous and discrete distributions;
and 2) implement these measurements without using histograms because they are not needed to
attain the cumulative distribution used for the Kolmogorov-Smirnov statistic and because
$c'$ is a metric (not a generalized metric) if obtained without using histograms,
as exposed in Section~\ref{sec:desc}.

% discrete vs continous
% are histograms really needed?

% use other distributions
% organize the scripts
% make a better presentation of the benchmark tables
% compare against other statistical distances

@@

% ------------------------------------------------------------------------
\begin{thebibliography}{99}
\fontsize{11}{0}\selectfont
\bibitem[Chicheportiche \& Bouchaud, 2012]{kolm}
Chicheportiche, R., \& Bouchaud, J. P. (2012). Weighted Kolmogorov-Smirnov test: Accounting for the tails. Physical Review E, 86(4), 041115.

\bibitem[DrV, 2016]{stack}
	DrV (2016). Python script to obtain the CDF from samples without doing the histogram.
		Stack Overflow.
		From \url{https://stackoverflow.com/questions/24788200/calculate-the-cumulative-distribution-function-cdf-in-python}

\bibitem[Fabbri, 2015]{gmaneLegacy}
	Fabbri, R. (2015). Gmane (legacy): a toolbox for analyzing topology and text of email interaction networks derived from the Gmane database. Github repositories. From \url{https://github.com/ttm/gmaneLegacy}

\bibitem[Fabbri, 2015]{ksstats}
	Fabbri, R. (2015). Benchmark tables for a statistical distance metric derived from the Kolmogorov-Smirnov test. Github repositories. From \url{https://github.com/ttm/kolmogorov-smirnov/raw/master/paper.pdf}
\end{thebibliography}
% ------------------------------------------------------------------------
